---
title: "Data 607 TidyVerse"
author: "Ajay Arora"
date: "August 27, 2019"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
#Load the library ggplot2
#remove.packages("ggplot2") # Unisntall ggplot
#install.packages("ggplot2") # Install it again
#install.packages("tidyverse")
library(tidyverse)
#library(ggplot2) # Load the librarie (you have to do this one on each new session)

```

```{r}
#Get the data from college majors
allages <- read.csv(file="./college-majors/all-ages.csv", header=TRUE, sep=",")
head(allages)
```
# 1.Initial ggplot 

```{r}
# A simple histogram using ggplot on Employed field
library(ggplot2)
ggplot(allages, aes(x = allages$Employed)) 
```

A blank ggplot is drawn. Even though the x and y are specified, there are no points or lines in it. This is because, ggplot doesn't assume that you meant a scatterplot or a line chart to be drawn. I have only told ggplot what dataset to use and what columns should be used for X and Y axis. I haven't explicitly asked it to draw any points.

Also note that aes() function is used to specify the X and Y axes. That's because, any information that is part of the source dataframe has to be specified inside the aes() function.


# 2. How to Make a Simple Histogram Plot
Aesthetic Mapping

In ggplot land aesthetic means "something you can see". Examples include:

    position (i.e., on the x and y axes)
    color ("outside" color)
    fill ("inside" color)
    shape (of points)
    linetype
    size

Each type of geom accepts only a subset of all aesthetics-refer to the geom help pages to see what mappings each geom accepts. Aesthetic mappings are set with the aes() function.

```{r}
# A simple histogram using ggplot on Employed field
library(ggplot2)
ggplot(allages, aes(x = allages$Employed)) + geom_histogram()
```

# 3. A histogram with adjustment for binning. (bin width)
```{r}
# A simple histogram using ggplot on Employed field
library(ggplot2)
ggplot(allages, aes(x = allages$Employed)) + geom_histogram(stat = "bin", binwidth=4000)
```



# 4. Geometic objects with ggplots 
Geometric objects are the actual marks we put on a plot. Examples include:

    points (geom_point, for scatter plots, dot plots, etc)
    lines (geom_line, for time series, trend lines, etc)
    boxplot (geom_boxplot, for, well, boxplots!)

A plot must have at least one geom; there is no upper limit. You can add a geom to a plot using the + operator
```{r}
#A scatter plot of college major Category and Employed
ggplot(allages,
       aes(x=allages$Employed,
           y=allages$Major_category
           ))+
  geom_point()

```

# 5. Linear Model with ggplots
```{r}
#A scatter plot of college major Category and Employed
ggplot(allages,
       aes(x=allages$Employed,
           y=allages$Major_category
           ))+
  geom_point()+ geom_smooth(method="lm")

```


# 6. Introducing color with ggplots
```{r}
#A scatter plot of college major Category and Employed
ggplot(allages,
       aes(x=allages$Employed,
           y=allages$Major_category
           ))+
  geom_point(aes(color = allages$Major_category))

```

The previous example does not make use of Faceting.


# 7. Color specific categories with ggplots
```{r}
#A scatter plot of college major Category and Employed
ggplot(allages,
       aes(x=allages$Employed,
           y=allages$Major_category
           ))+
  geom_point(aes(color = allages$Major_category, size =  allages$Total))

```


# 8. Faceting with ggplots 
Faceting

    Faceting is ggplot2 parlance for small multiples
    The idea is to create separate graphs for subsets of data
    ggplot2 offers two functions for creating small multiples:
        facet_wrap(): define subsets as the levels of a single grouping variable
        facet_grid(): define subsets as the crossing of two grouping variables
    Facilitates comparison among plots, not just of geoms within a plot


# 9. Different shapes with ggplots and Faceting
```{r}

ggplot(allages,
       aes(x=allages$Employed,
           y=allages$Major_category
           , group = 1))+
   geom_line() +
   facet_wrap(~allages$Major_category, ncol = 16) 

```


# 10. Themes with ggplots
Themes

The ggplot2 theme system handles non-data plot elements such as

    Axis labels
    Plot background
    Facet label backround
    Legend appearance

Built-in themes include:

    theme_gray() (default)
    theme_bw()
    theme_classc()


```{r}

ggplot(allages,
       aes(x=allages$Employed,
           y=allages$Major_category
           , group = 1))+
   geom_line() +
   facet_wrap(~allages$Major_category, ncol = 16) +  theme_linedraw()

```









References: 
http://r-statistics.co/Complete-Ggplot2-Tutorial-Part1-With-R-Code.html
https://tutorials.iq.harvard.edu/R/Rgraphics/Rgraphics.html
https://ggplot2.tidyverse.org/


